---
title: 'STAT 5014: Homework Two'
author: "Bobby Soule"
date: "9/12/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(lubridate)
library(data.table)
setwd("/Users/bobbysoule/Documents/College/Graduate/STAT_5014/STAT_5014_homework/02_data_munging_summarizing_R_git")
```

## Problem 4

I think using Git will primarily help me by providing an easy-to-use backup system.  I will not have to worry about losing files; and if I make a mistake while editing a project, then I can simply revert back to the previous version.  Additionally, I think it will be very helpful in any future scenarios where I collaborate on a project.

## Problem 5

## Part A: Sensory Data

```{r}
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensory_data <- read.table(url, header=FALSE, skip=1, fill=TRUE, stringsAsFactors=F) %>%
  tbl_df

sensory_data <- sensory_data[-1,]
colnames(sensory_data) <- c("Item", "V1", "V2", "V3", "V4", "V5")

sensory_data_NArows <- filter(.data=sensory_data, is.na(V5)) %>%
  select(V1=Item, V2=V1, V3=V2, V4=V3, V5=V4) %>%
  mutate(Item=rep(as.character(1:(nrow(sensory_data)/3)), each=2), 
         V1=as.numeric(V1)) %>%
  select(Item, V1:V5)

sensory_data <- filter(.data=sensory_data, !is.na(V5)) %>%
  bind_rows(sensory_data_NArows)
colnames(sensory_data)[2:6] <- paste("Operator", 1:5, sep="_")

sensory_final <- gather(data=sensory_data, key=Operator, value=Measure,
                        Operator_1:Operator_5) %>%
  mutate(Operator = gsub("Operator_", "", Operator)) %>%
  arrange(Item)
```

```{r}
    knitr::kable(summary(sensory_final), caption="Sensory Data Summary")
```

## Part B: Gold Medal Olympic Long Jump Distances (Men)

```{r}
url2 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
ljump_data <- read.table(url2, header=FALSE, skip=1, fill=TRUE, stringsAsFactors=FALSE) %>%
  tbl_df

colnames(ljump_data) <- rep(c("Year", "Distance"), times=4)
ljump_final <- rbind(ljump_data[,1:2], ljump_data[,3:4],ljump_data[,5:6], ljump_data[,7:8]) %>%
  filter(!is.na(Year)) %>%
  mutate(Year = Year+1900)
```

```{r}
    knitr::kable(summary(ljump_final), caption="Long Jump Data Summary")
```

## Part C: Brain Weight vs Body Weight

```{r}
url3 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
brain_body_data <- read.table(url3, header=FALSE, skip=1, fill=TRUE, stringsAsFactors=FALSE) %>%
  tbl_df

colnames(brain_body_data) <- rep(c("Body_Wt", "Brain_Wt"), times=3)
brain_body_final <- rbind(brain_body_data[,1:2], brain_body_data[,3:4], brain_body_data[,5:6]) %>%
  filter(!is.na(Body_Wt))
```

```{r}
    knitr::kable(summary(brain_body_final), caption="Brain & Body Weight Data Summary")
```
  
## Part D: Tomato Yield vs. Variety & Planting Density

```{r}
url4 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tom_data <- read.table(url4, header=FALSE, skip=2, fill=TRUE,
                       stringsAsFactors=FALSE, comment.char = "")
colnames(tom_data) <- c("Variety", "1000", "2000", "3000")
tom_data[1,1] <- "Ife_1"
tom_data[2,2] <- gsub("10.1,", "10.1", tom_data[2,2])

tom_final <- tom_data %>%
  gather(key=Density, value=Yield, -Variety) %>%
  separate(Yield, into=paste("Yield", 1:3, sep="_"), sep=",") %>%
  gather(key=Replicate, value=Yield, Yield_1:Yield_3) %>%
  mutate(Replicate = gsub("Yield_", "", Replicate), Yield=as.numeric(Yield)) %>%
  arrange(Variety, Density)
```

```{r}
    knitr::kable(summary(tom_final), caption="Tomato Yield Data summary")
```

## Problem 6

```{r}

# Path to data
library(swirl)
datapath <- file.path(path.package('swirl'), 'Courses', 'R_Programming_E',
                      'Looking_at_Data', 'plant-data.txt')
plants <- read.csv(datapath, strip.white=TRUE, na.strings="") %>%
          tbl_df
detach(package:swirl)

# Remove annoying columns
cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol')
plants <- plants[, !(names(plants) %in% cols2rm)]

# Make names pretty
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period',
                   'Foliage_Color', 'pH_Min', 'pH_Max','Precip_Min',
                   'Precip_Max','Shade_Tolerance', 'Temp_Min_F')

plants_final <- plants %>%
                mutate(pH_Mid = (pH_Max+pH_Min)/2) %>%
                select(Scientific_Name, Foliage_Color, pH_Mid) %>%
                filter(!is.na(Foliage_Color) | !is.na(pH_Mid))

plot(plants_final$Foliage_Color, plants_final$pH_Mid)
plants_reg <- lm(formula = pH_Mid ~ Foliage_Color, data = plants_final)
plants_anova <- anova(plants_reg)
```

```{r}
knitr::kable(summary(plants_reg)$coef, caption="Regression Coefficients")
knitr::kable(plants_anova, caption="ANOVA Table")
```

## Problem 7

```{r}
#defect code, defect comment
Car_Gebreken <- fread(input = "Open_Data_RDW__Gebreken.csv",
                             header = T, select=c(1,6), showProgress=F) %>% tbl_df
colnames(Car_Gebreken) <- c("Code", "Comment")
#license plate, inspection date, defect code
Car_Geconstat <- fread(input = "Open_Data_RDW__Geconstateerde_Gebreken.csv",
                              header=T, select=c(1,3,5),showProgress=F) %>% tbl_df
colnames(Car_Geconstat) <- c("Plate", "Inspect_Date", "Code")
Car_Geconstat <- mutate(Car_Geconstat, Inspect_Date = ymd(Inspect_Date)) %>%
                 filter(year(Inspect_Date) == 2017)
#license plate, make, model
Car_Person <- fread(input = "Personenauto_basisdata.csv",
                           header=T, showProgress = F, select = c(1,3,4)) %>% tbl_df
colnames(Car_Person) <- c("Plate", "Make", "Model")

Car_Final <- full_join(x = Car_Geconstat, y = Car_Person, by = "Plate") %>%
             full_join(y = Car_Gebreken, by = "Code")
Car_Final <- Car_Final[complete.cases(Car_Final),]

unique_make <- n_distinct(Car_Final$Make)
unique_make
unique_model <- n_distinct(Car_Final$Model)
unique_model

Car_byCode <- group_by(Car_Final, Code, Comment)
Freq_Defects <- summarize(Car_byCode, Count=n()) %>%
                arrange(desc(Count)) %>%
                head(5)
top5_defects <- Freq_Defects$Code

Car_byCodeMakeModel <- group_by(Car_Final, Code, Make, Model)
Freq_Defects2 <- summarize(Car_byCodeMakeModel, Count=n()) %>%
                 filter(Code %in% top5_defects) %>%
                 arrange(desc(Count))

for (code in top5_defects) {
  assign(code,
         filter(Freq_Defects2, Code==code) %>%
         arrange(desc(Count)) %>%
         head(5) %>% ungroup %>%
         mutate(Tot_Count = sum(Count))
         )
}

Freq_Defects_Final <- get(top5_defects[1])
for (i in 2:length(top5_defects)) {
  Freq_Defects_Final <- rbind(Freq_Defects_Final, get(top5_defects[i]))
}

tot_counts <- Freq_Defects$Count
names(tot_counts) <- Freq_Defects$Code
Freq_Defects_Final <- mutate(Freq_Defects_Final, Tot_Count = tot_counts[Code])

knitr::kable(Freq_Defects_Final, caption="Top Five Defects")

cars_reg <- lm(formula = Count ~ Make, data = Freq_Defects2)
cars_anova <- anova(cars_reg)

knitr::kable(summary(cars_reg)$coef, caption="Regression Coefficients")
knitr::kable(cars_anova, caption="ANOVA Table")

```
